Field and experimental data together with computational

### 2022-03-08 manuscript_PLOS/manuscript_PLOS.tex
Updated manuscript tp the PLOS Comp. Biol. Updated figures with new figures.

%% Author_tex.tex
%% V1.0
%% 2012/13/12
%% developed by Techset
%%
%% This file describes the coding for rsproca.cls

\documentclass[]{rsos}%%%%where rsos is the template name

\usepackage{lineno}
\linenumbers

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

@@

@@

%%%% *** Do not adjust lengths that control margins, column widths, etc. ***

%%%%%%%%%%% Defining Enunciations  %%%%%%%%%%%
\newtheorem{theorem}{\bf Theorem}[section]
\newtheorem{condition}{\bf Condition}[section]
\newtheorem{corollary}{\bf Corollary}[section]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%% Article title to be placed here
\title{Field and experimental data together with computational models reveal how cleaner fish adjust decisions in a biological market}

\author{
Andrés E. Quiñones$^{1}$,
Zegni Triki$^{2}$,
Redouan Bshary$^{1}$}

\address{
  $^{1}$Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland\\
  $^{2}$Department of Zoology, Stockholm University, Stockholm, Sweden}
%%%% Subject entries to be placed here %%%%
\subject{
Behavioural ecology,
Cognitive ecology,
Animal behaviour}

%%%% Keyword entries to be placed here %%%%
\keywords{
learning,
cleaners,
behaviour,
bayesian statisitics,
mutualism}

%%%% Insert corresponding author and its email address}
\corres{
  AE. Quiñones\\
  e-mail: \href{mailto:andreseqp@gmail.com}{\nolinkurl{andreseqp@gmail.com}}
}

%%%% Abstract text to be placed here %%%%%%%%%%%%
\begin{abstract}
While it is generally straightforward to quantify individual performance in cognitive experiments, identifying the underlying cognitive processes remains a major challenge. Often, different mechanistic underpinnings yield similar performances, and Lloyd Morgan's cannon warrants acceptance of the simpler explanation. Alternatively, when the different mechanisms interact with environmental conditions, variation in performance across environments might allow to statistically infer the mechanism responsible. We illustrate this point by fitting computational models to experimental data on performance by wild-caught cleaner fish Labroides dimidiatus in an ephemeral reward task, as well as cleaner and client fish densities from the locations of capture. Using Bayesian statistics to fit the model parameters to performance data revealed that cleaner fish most likely estimate future consequences of an action, while it appears unlikely that the removal of the ephemeral reward acts as psychological punishment (negative reinforcement). Incorporating future consequences also yields performances that can be considered the result of locally optimal decision-rules, in contrast to the negative reinforcement mechanism. We argue that the combination of computational models with data is a powerful tool to infer the mechanistic underpinnings of cognitive performance.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%

\providecommand{\EndFirstPage}{%
}

\maketitle

@@

Often alternative cognitive mechanisms yield similar behaviour and/or cognitive
performances. This poses a problem for disentangling the mechanistic
underpinnings of behaviour. This is particularly clear in research aimed at
discovering between species variation in \emph{higher} cognitive abilities; or
in other words, research on whether non-human animals show cognitive
abilities believed to be uniquely human. For instance, when researchers
try to find \emph{mental time travel-like} behaviour, they usually come up with
clever experiments to show the behaviour displayed requires inferences made
through past events\citep{dally_Foodcaching_2006}. However, they often face the
challenge of alternative scenarios where simpler explanations,
like classic associative learning, can bring about the
observed behavioural outcome \citep{suddendorf_Evolution_2007}. Similarly, attempts
to demonstrate the presence of \emph{theory of mind} in non-human animals
face objections justified by alternative mechanisms underpinning
similar behavioural results \citep{heyes_Theory_1998}.
Such controversies are usually settled by using the principle
of parsimony and its cognitive version,
Lloyd Morgan's cannon, which states that the simpler explanation (mechanism)
should be accepted. Ideally, however, alternative hypotheses should be evaluated
in light of their explanatory power. One potential dimension that could allow
the variation necessary among distinct cognitive mechanisms is environmental
variation. Controlled laboratory conditions often mask between and within
species variation that arises among different environments. Thus, if
predictions from the alternative mechanisms depend on environmental
conditions, it is possible to assess their relative explanatory power
by performing cognitive test in the context of
different environmental conditions.

From an evolutionary perspective, mechanisms are likely selected because
of how they allow individuals to respond to environmental variation.
For example, biological market theory predicts that the exchange rate of goods
and/or services traded between cooperative partners adjusts to the
law of supply and demand, when individuals have some degree of partner choice
\citep{noe_Biological_1995}. Supply and demand conditions, which typically
depend on the abundance of the species involved, certainly vary in time
and space. Therefore, natural selection should favour the ability
to flexibly adjust decisions and behavioural output to
current market conditions. Indeed, such adjustments have been repeatedly
documented \citep{axen_Signalling_1996}. In animals, an obvious general candidate
mechanism for the strategic adjustment is the cognitive machinery. However,
it is not clear which cognitive mechanisms allow individuals to adjust their
behaviour to the varying conditions. More specifically the question arises
to what extent mechanisms beyond basic associative learning may be involved.

One example of strategic adjustment in a biological market is the marine
cleaning mutualism involving the cleaner fish \emph{Labroides dimidiatus} and
`client' fish. Client fish seek cleaner fish services at their territory
(so-called ``cleaning station'') and offer themselves as food patches
to get their ectoparasites removed, which provides cleaners
with food and clients with improved health \citep{waldie_LongTerm_2011, ros_Does_2011, triki_Effects_2016, demaire_Reduced_2020}.
Given the capacity of some client fish to swim larger distances and
access multiple cleaning stations while others access the only cleaning
station in their territory, it is crucial to categorize clients as
either ``visitors'' or ``residents'', respectively. During cleaning interactions,
a cleaner fish often face a choice between a visitor and a resident client
seeking its cleaning services simultaneously. Visitors have the option
to switch to another cleaner fish if being made to wait, while residents
must wait for inspection. Indeed,
visitors have been observed to use their partner choice option in that
way \citep{bshary_Choosy_2002}, which may explain why cleaners give
visitors service priority in a field study in the Red Sea
\citep{bshary_Cleaner_2001}.

When researchers aimed at testing wild-caught cleaner fish, as
well as individuals from other species, abilities
to prioritize visitor over resident clients in a lab-based paradigm,
they used Plexiglas plates of different colours and shapes offering the same
amount of food as surrogates for visitor and resident clients.
One plate is made to behave like a resident, i.e.~it remains until
the cleaner fish has eaten off it. The other plate was made to behave
like a visitor, i.e.~it is removed if not inspected first.
This laboratory paradigm originally inspired by the cleaning market
of cleaner fish is more generally referred as the ephemeral reward task.
Individuals are assessed in their capacity to develop a preference for the
ephemeral option, which gives them access to two reward units.
Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food \citep{bshary_Asymmetric_2002}.
Furthermore, adult cleaner fish outperformed various primates as well as rats and
pigeons in this original version of the ephemeral reward task
\citep{salwiczek_Adult_2012, zentall_Early_2017}; yet the African grey
parrots solve this task as well \citep{pepperberg_Can_2014}.
This between variation in performance, and primates' known cognitive abilities
\(\color{red}{\text{ref}}\), shows that performance in the task
cannot be readily explained by classifying species as having high and
low cognitive abilities. Instead, this between species variation suggests
that solving the ephemeral reward task requires a set of cognitive
mechanisms that are not present in all the tested taxa, or were not used
under the experimental set-up.

Recent research on the cognitive abilities needed to solve the ephemeral
reward task has used a broad approach as proposed by Shettleworth
\citep{shettleworth_Cognition_2009}, who defined cognition as including
all ways in which animals acquire information
through the senses, process, retain and decide to act on it.
Indeed, perception of relevant cues is of major importance
for individual performance. For example, some animals find relevant
information in the plates \citep{wismer_Cuebased_2019} while others find
it in the food \citep{pretot_Comparative_2021, pretot_Comparing_2016}.
However, identifying cues as salient is not the only challenge
of the ephemeral reward task, as revealed by proximate learning models.
Such models allow varying the cognitive tool kit and evaluating
which minimal kit is necessary to solve
the task at hand (e.g. \citep{dubois_Model_2021}). Applied to the ephemeral
reward task, learning models showed that basic reinforcement learning does
not suffice to solve the ephemeral reward task \citep{prat_Modelling_2022, quinones_Reinforcement_2019}.
This is particularly so when models assume the more complex natural
situation in which cleaner fish face not only resident-visitor pairs but also
visitor-visitor and resident-resident pairs, as well a resident alone or
a visitor alone. To be able to give visitors priority over residents,
cleaners need to be able to assess a client's value separately for the
three possible scenarios (alone, paired with a fish with the
same strategic option, paired with a fish with the alternative strategic option)
\citep{quinones_Reinforcement_2019}. The ability to distinguish and value
one stimulus differently alone from compound versions of it has been termed
configurational learning, chunking,
or segmentation (see references in \citep{prat_Modelling_2022}).

In addition to configurational learning, cleaners also need to account for
the future consequences of current decisions. In the model by
Quiñones \emph{et al.} \citep{quinones_Reinforcement_2019}, this could be achieved in
two non-mutually exclusive ways: through low temporal discounting of
future effects, also termed `chaining' \citep{enquist_Power_2016}; and/or
through perceiving a visitor client leaving as psychological punishment
(i.e.~as a negative reinforcer). Low temporal discounting is when
individuals include in their valuation of an action the reward effects
that this will have in the future. This is done by combining in a single
valuation the reward obtained in the current time with all the reward that
comes after, discounting for how far in the future reward
is accrued. `Chaining' the reward of these different time steps allows
individuals to take actions that increase the long-term reward at the sacrifice
of short term considerations \citep{enquist_Power_2016}. Even though, `chaining'
can be readily implemented computationally in learning models
\citep{enquist_Power_2016, sutton_Reinforcement_2018}, cognitively it seems to be
a complex adaptation \citep{suddendorf_Evolution_2007}. On the other hand,
using client behaviour as a negative reinforcer is, in principle,
easier to implement. Negative reinforcement is part of the ubiquitous
learning mechanism termed operant conditioning \citep{thorndike_Animal_1898, skinner_Behavior_1938}. Thus, the standard logic of
Lloyd Morgan's cannon demands that operant conditioning as the simpler
explanation is to be accepted by default. Ideally, however, the two
mechanisms should be evaluated in light of how well
they explain the available data. Note that different fields interested in
cognition and decision making use different words to refer to negative
reinforcers \citep{quinones_Reinforcement_2019, sutton_Reinforcement_2018}.
Here, for the sake of simplicity and clarity, we will use the word
`penalty' to refer to this mechanism which includes a negative reinforcer.

Over the last decade, our team tested over a hundred wild-caught
cleaner fish in the exact same paradigm of the ephemeral reward
task \citep{salwiczek_Adult_2012, wismer_Variation_2014, triki_Decrease_2018, triki_Biological_2019, triki_Brain_2020}. These fish often come from
different reef locations around the study field site at Lizard Island,
Great Barrier Reef in Australia. Further investigation of the local
eco-sociological conditions revealed that cleaner and client
fish population densities have a substantial impact on cleaner fish
performance in the task. Cleaner fish from reef sites with relatively
low densities were more likely to fail at solving the task \citep{triki_Biological_2019, triki_Decrease_2018, wismer_Variation_2014}. The explanation for this pattern
is twofold. First, low cleaner fish population densities imply
low supply and high demand for cleaning services.
Under such conditions, there are fewer occasions under
which failing to give priority to a visitor translates to an empty
cleaning station. Second and also related to the state of the cleaning market,
visitor clients at low cleaner fish population densities
are less likely to exert their partner choice options
and are hence more likely to wait for inspection if not given cleaning
priority \citep{triki_Brain_2020}. In the absence of partner choice
cleaner fish should not give priority to visitors any more.

@@

The model consists of a set of individual-based simulations where
individuals, representing cleaner fish, face a series of choices
between two options, which simulate the natural conditions
of the cleaning market. Individuals experience a series
of discrete time points in which they face different `states',
defined by the number and category of client fish (visitor or residents)
inviting for cleaning services. There are six possible states:
zero clients, one resident, one visitor,
resident-resident, visitor-visitor, and resident-visitor.
The probability of each state is largely determined by the relative
abundance of cleaner fish, residents and visitors, but to some degree by
cleaner fish choices when it faces the resident-visitor combination.
This is because residents are willing to queue for cleaning service;
while visitors leave the queue (with a certain probability) when made to wait.
Individuals obtain a fixed reward from cleaning a client fish
regardless of the category. Every time individuals face and make a choice
they update the probability of making that same choice. The update is
based on the difference between the expected reward and the obtained
reward (prediction error) \citep{sutton_Reinforcement_2018, rescorla_Theory_1972}. Given the new information, the update is carried
in the direction that leads to more reward being obtained, given
the new information. In the long run, the probability of choosing a visitor
over a resident converges in the model. To which probability the model
will converge depends on the relative abundance of cleaners, visitors
and residents; as well as on the probability of visitors leaving the
cleaning station when made to wait. Further details of the model
implementation can be found in Quiñones \emph{et al} \citep{quinones_Reinforcement_2019}.

@@

The aim of the analyses is to fit the key model parameters
\(\gamma\) and \(\eta\), to the empirical data from Triki \emph{et al.}
\citep{triki_Biological_2019, triki_Brain_2020} to test whether each or
a combination of these effects is a better explanation for the pattern
seen in the data. We used the ecological variables: cleaners, visitor clients,
resident clients abundances and visitor clients leaving probability,\\
as input to the models. As the response variable, we used cleaners'
preference for visitors over residents in the ephemeral reward task.
Finally, we used the preference for visitors resulting from the
model simulations as the prediction for the response variable.
We kept all other parameter values used for the model simulations constant,
see Table \ref{tab:param}.

@@

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/post-1}

}

>>>>

% Template for PLoS
% Version 3.5 March 2018
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended
% to minimize problems and delays during our production
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% Once your paper is accepted for publication,
% PLEASE REMOVE ALL TRACKED CHANGES in this file
% and leave only the final text of your manuscript.
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file.
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission.
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.
%
% Do not include text that is not math in the math environment.
%
% Please add line breaks to long display equations when possible in order to fit size of the column.
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
% \usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace}
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
% \bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

@@

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

\usepackage{forarray}
\usepackage{xstring}
\newcommand{\getIndex}[2]{
  \ForEach{,}{\IfEq{#1}{\thislevelitem}{\number\thislevelcount\ExitForEach}{}}{#2}
}

\setcounter{secnumdepth}{0}

\newcommand{\getAff}[1]{
  \getIndex{#1}{Institute of Biology, University of Neuchatel,Department
of Zoology, Stockholm University}
}

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Field and experimental data together with computational
models reveal how cleaner fish adjust decisions in a biological
market} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Andrés E. Quiñones\textsuperscript{\getAff{Institute of Biology,
University of Neuchatel}}\textsuperscript{*},
Zegni Triki\textsuperscript{\getAff{Department of Zoology, Stockholm
University}},
Redouan Bshary\textsuperscript{\getAff{Institute of Biology, University
of Neuchatel}}\\
\bigskip
\textbf{\getAff{Institute of Biology, University of Neuchatel}}Institute
of Biology, Rue Emile-Argand 11, 2000-CH Neuchatel, Switzerland\\
\textbf{\getAff{Department of Zoology, Stockholm University}}Department
of Zoology, Svante Arrheniusväg 18 B, 106 91 Stockholm, Sweden,\\
\bigskip
* Corresponding author: andresqp@gmail.com\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
While it is generally straightforward to quantify individual performance
in cognitive experiments, identifying the underlying cognitive processes
remains a major challenge. Often, different mechanistic underpinnings
yield similar performances, and Lloyd Morgan's cannon warrants
acceptance of the simpler explanation. Alternatively, when the different
mechanisms interact with environmental conditions, variation in
performance across environments might allow to statistically infer the
mechanism responsible. We illustrate this point by fitting computational
models to experimental data on performance by wild-caught cleaner fish
\emph{Labroides dimidiatus} in an ephemeral reward task, as well as
cleaner and client fish densities from the locations of capture. Using
Bayesian statistics to fit the model parameters to performance data
revealed that cleaner fish most likely estimate future consequences of
an action, while it appears unlikely that the removal of the ephemeral
reward acts as psychological punishment (negative reinforcement).
Incorporating future consequences also yields performances that can be
considered the result of locally optimal decision-rules, in contrast to
the negative reinforcement mechanism. We argue that the combination of
computational models with data is a powerful tool to infer the
mechanistic underpinnings of cognitive performance.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step.
% Author Summary not valid for PLOS ONE submissions.
\section*{Author summary}
Performance in behavioural experiments is often used to assess the
cognitive abilities of animals. However, animals can get to the same
outcome in alternative ways. Thus, the outcome of the experiments does
not tell us how animals achieve their performance. In order to overcome
this limitation, we used a set of computational models, which provide
predictions on how alternative cognitive mechanisms perform in the face
of varying environmental conditions, and compare their predictions to
performance data from individuals that experienced different conditions
throughout their life. Our study system is the cleaner fish
\emph{Labroides dimidiatus}, a coral reef fish that feeds off
ectoparasites and dead tissue from the skin of other reef fishes.
Cleaner fish often have to choose among alternative clients seeking
their cleaning service, and the optimal choice is often hard to achieve.
Our performance experiments mimic the natural choices the cleaner fish
make. The combination of experiments and computational models points to
cleaner fish abilities to account for the future effect of their choices
by estimating the long term expected value of those choices. Estimating
long term value is a mechanism involved in human foresight.

\linenumbers

% Use "Eq" instead of "Equation" for equation citations.
\emph{Text based on plos sample manuscript, see
\url{https://journals.plos.org/ploscompbiol/s/latex}}

@@

Often alternative cognitive mechanisms yield similar behaviour and/or
cognitive performances. This poses a problem for disentangling the
mechanistic underpinnings of behaviour. This is particularly clear in
research aimed at discovering between species variation in \emph{higher}
cognitive abilities; or in other words, research on whether non-human
animals show cognitive abilities believed to be uniquely human. For
instance, when researchers try to find \emph{mental time travel-like}
behaviour, they usually come up with clever experiments to show the
behaviour displayed requires inferences made through past events{[}1{]}.
However, they often face the challenge of alternative scenarios where
simpler explanations, like classic associative learning, can bring about
the observed behavioural outcome {[}2{]}. Similarly, attempts to
demonstrate the presence of \emph{theory of mind} in non-human animals
face objections justified by alternative mechanisms underpinning similar
behavioural results {[}3{]}. Such controversies are usually settled by
using the principle of parsimony and its cognitive version, Lloyd
Morgan's cannon, which states that the simpler explanation (mechanism)
should be accepted. Ideally, however, alternative hypotheses should be
evaluated in light of their explanatory power. One potential dimension
that could allow the variation necessary among distinct cognitive
mechanisms is environmental variation. Controlled laboratory conditions
often mask between and within species variation that arises among
different environments. Thus, if predictions from the alternative
mechanisms depend on environmental conditions, it is possible to assess
their relative explanatory power by performing cognitive test in the
context of different environmental conditions.

From an evolutionary perspective, mechanisms are likely selected because
of how they allow individuals to respond to environmental variation. For
example, biological market theory predicts that the exchange rate of
goods and/or services traded between cooperative partners adjusts to the
law of supply and demand, when individuals have some degree of partner
choice {[}4{]}. Supply and demand conditions, which typically depend on
the abundance of the species involved, certainly vary in time and space.
Therefore, natural selection should favour the ability to flexibly
adjust decisions and behavioural output to current market conditions.
Indeed, such adjustments have been repeatedly documented {[}5{]}. In
animals, an obvious general candidate mechanism for the strategic
adjustment is the cognitive machinery. However, it is not clear which
cognitive mechanisms allow individuals to adjust their behaviour to the
varying conditions. More specifically the question arises to what extent
mechanisms beyond basic associative learning may be involved.

One example of strategic adjustment in a biological market is the marine
cleaning mutualism involving the cleaner fish \emph{Labroides
dimidiatus} and `client' fish. Client fish seek cleaner fish services at
their territory (so-called ``cleaning station'') and offer themselves as
food patches to get their ectoparasites removed, which provides cleaners
with food and clients with improved health {[}6--9{]}. Given the
capacity of some client fish to swim larger distances and access
multiple cleaning stations while others access the only cleaning station
in their territory, it is crucial to categorize clients as either
``visitors'' or ``residents,'' respectively. During cleaning
interactions, a cleaner fish often face a choice between a visitor and a
resident client seeking its cleaning services simultaneously. Visitors
have the option to switch to another cleaner fish if being made to wait,
while residents must wait for inspection. Indeed, visitors have been
observed to use their partner choice option in that way {[}10{]}, which
may explain why cleaners give visitors service priority in a field study
in the Red Sea {[}11{]}.

When researchers aimed at testing wild-caught cleaner fish, as well as
individuals from other species, abilities to prioritize visitor over
resident clients in a lab-based paradigm, they used Plexiglas plates of
different colours and shapes offering the same amount of food as
surrogates for visitor and resident clients. One plate is made to behave
like a resident, i.e.~it remains until the cleaner fish has eaten off
it. The other plate was made to behave like a visitor, i.e.~it is
removed if not inspected first. This laboratory paradigm originally
inspired by the cleaning market of cleaner fish is more generally
referred as the ephemeral reward task. Individuals are assessed in their
capacity to develop a preference for the ephemeral option, which gives
them access to two reward units. Cleaners learned to prefer the visitor
plate and hence obtained the double amount of food {[}12{]}.
Furthermore, adult cleaner fish outperformed various primates as well as
rats and pigeons in this original version of the ephemeral reward task
{[}13,14{]}; yet the African grey parrots solve this task as well
{[}15{]}. This between variation in performance, and primates' known
cognitive abilities \(\color{red}{\text{ref}}\), shows that performance
in the task cannot be readily explained by classifying species as having
high and low cognitive abilities. Instead, this between species
variation suggests that solving the ephemeral reward task requires a set
of cognitive mechanisms that are not present in all the tested taxa, or
were not used under the experimental set-up.

Recent research on the cognitive abilities needed to solve the ephemeral
reward task has used a broad approach as proposed by Shettleworth
{[}16{]}, who defined cognition as including all ways in which animals
acquire information through the senses, process, retain and decide to
act on it. Indeed, perception of relevant cues is of major importance
for individual performance. For example, some animals find relevant
information in the plates {[}17{]} while others find it in the food
{[}18,19{]}. However, identifying cues as salient is not the only
challenge of the ephemeral reward task, as revealed by proximate
learning models. Such models allow varying the cognitive tool kit and
evaluating which minimal kit is necessary to solve the task at hand
(e.g. {[}20{]}). Applied to the ephemeral reward task, learning models
showed that basic reinforcement learning does not suffice to solve the
ephemeral reward task {[}21,22{]}. This is particularly so when models
assume the more complex natural situation in which cleaner fish face not
only resident-visitor pairs but also visitor-visitor and
resident-resident pairs, as well a resident alone or a visitor alone. To
be able to give visitors priority over residents, cleaners need to be
able to assess a client's value separately for the three possible
scenarios (alone, paired with a fish with the same strategic option,
paired with a fish with the alternative strategic option) {[}22{]}. The
ability to distinguish and value one stimulus differently alone from
compound versions of it has been termed configurational learning,
chunking, or segmentation (see references in {[}21{]}).

In addition to configurational learning, cleaners also need to account
for the future consequences of current decisions. In the model by
Quiñones \emph{et al.} {[}22{]}, this could be achieved in two
non-mutually exclusive ways: through low temporal discounting of future
effects, also termed `chaining' {[}23{]}; and/or through perceiving a
visitor client leaving as psychological punishment (i.e.~as a negative
reinforcer). Low temporal discounting is when individuals include in
their valuation of an action the reward effects that this will have in
the future. This is done by combining in a single valuation the reward
obtained in the current time with all the reward that comes after,
discounting for how far in the future reward is accrued. `Chaining' the
reward of these different time steps allows individuals to take actions
that increase the long-term reward at the sacrifice of short term
considerations {[}23{]}. Even though, `chaining' can be readily
implemented computationally in learning models {[}23,24{]}, cognitively
it seems to be a complex adaptation {[}2{]}. On the other hand, using
client behaviour as a negative reinforcer is, in principle, easier to
implement. Negative reinforcement is part of the ubiquitous learning
mechanism termed operant conditioning {[}25,26{]}. Thus, the standard
logic of Lloyd Morgan's cannon demands that operant conditioning as the
simpler explanation is to be accepted by default. Ideally, however, the
two mechanisms should be evaluated in light of how well they explain the
available data. Note that different fields interested in cognition and
decision making use different words to refer to negative reinforcers
{[}22,24{]}. Here, for the sake of simplicity and clarity, we will use
the word `penalty' to refer to this mechanism which includes a negative
reinforcer.

Over the last decade, our team tested over a hundred wild-caught cleaner
fish in the exact same paradigm of the ephemeral reward task
{[}13,27--30{]}. These fish often come from different reef locations
around the study field site at Lizard Island, Great Barrier Reef in
Australia. Further investigation of the local eco-sociological
conditions revealed that cleaner and client fish population densities
have a substantial impact on cleaner fish performance in the task.
Cleaner fish from reef sites with relatively low densities were more
likely to fail at solving the task {[}27--29{]}. The explanation for
this pattern is twofold. First, low cleaner fish population densities
imply low supply and high demand for cleaning services. Under such
conditions, there are fewer occasions under which failing to give
priority to a visitor translates to an empty cleaning station. Second
and also related to the state of the cleaning market, visitor clients at
low cleaner fish population densities are less likely to exert their
partner choice options and are hence more likely to wait for inspection
if not given cleaning priority {[}30{]}. In the absence of partner
choice cleaner fish should not give priority to visitors any more.

@@

The model consists of a set of individual-based simulations where
individuals, representing cleaner fish, face a series of choices between
two options, which simulate the natural conditions of the cleaning
market. Individuals experience a series of discrete time points in which
they face different `states,' defined by the number and category of
client fish (visitor or residents) inviting for cleaning services. There
are six possible states: zero clients, one resident, one visitor,
resident-resident, visitor-visitor, and resident-visitor. The
probability of each state is largely determined by the relative
abundance of cleaner fish, residents and visitors, but to some degree by
cleaner fish choices when it faces the resident-visitor combination.
This is because residents are willing to queue for cleaning service;
while visitors leave the queue (with a certain probability) when made to
wait. Individuals obtain a fixed reward from cleaning a client fish
regardless of the category. Every time individuals face and make a
choice they update the probability of making that same choice. The
update is based on the difference between the expected reward and the
obtained reward (prediction error) {[}24,31{]}. Given the new
information, the update is carried in the direction that leads to more
reward being obtained, given the new information. In the long run, the
probability of choosing a visitor over a resident converges in the
model. To which probability the model will converge depends on the
relative abundance of cleaners, visitors and residents; as well as on
the probability of visitors leaving the cleaning station when made to
wait. Further details of the model implementation can be found in
Quiñones \emph{et al} {[}22{]}.

@@

The aim of the analyses is to fit the key model parameters \(\gamma\)
and \(\eta\), to the empirical data from Triki \emph{et al.} {[}29,30{]}
to test whether each or a combination of these effects is a better
explanation for the pattern seen in the data. We used the ecological
variables: cleaners, visitor clients, resident clients abundances and
visitor clients leaving probability,\\
as input to the models. As the response variable, we used cleaners'
preference for visitors over residents in the ephemeral reward task.
Finally, we used the preference for visitors resulting from the model
simulations as the prediction for the response variable. We kept all
other parameter values used for the model simulations constant, see
Table \ref{tab:param}.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_PLOS_files/figure-latex/post-1} \caption{Posterior distributions for parameter values $\gamma$ and $\eta$ for the three models (Full model, chaining only and penalty only). We show the kernel density estimates, below the mode (black dot) and the 65 (light blue shade) and 95\% (grey shade)  highest posterior density interval for the two parameters. On the top, panels a and b show posterior distributions for a full model, including chaining ($\gamma$) and penalty ($\eta$). Panel c, shows the $\gamma$ estimate from model with only chaining. Panel d shows the $\eta$ estimate from a model with only penalty. Panel e, shows a measure of fit for all models, namely the distribution of pseudo-$R^2$ obtained from sampling the posterior distribution of parameter values.}\label{fig:post}
\end{figure}
